---
title: "Lab2"
author: "Yangzom Dolma"
data: "09/04/2020"
output: 
  html_document: 
    toc: yes
    toc_float: yes 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```
## Task 2

```{r}
epagas = read.csv("EPAGAS.csv")
head(epagas, 6)
```
## Task 3

Make the object mpg, the number of miles per gallon vector

```{r}
mpg = epagas$ï..MPG
head(mpg)
```
Transform the mpg variable to z and verify these results.
```{r}
z= (mpg - mean(mpg))/sd(mpg)
head(z)
print(paste0("mean = ", round(mean(mpg), digits = 4)), quote = FALSE)
print(paste0("variance = ", var(z)), quote = FALSE)
print(paste0("standard deviation = ", sd(z)), quote = FALSE)
```
Using z, the values of mpg that are possible outliers are: 
```{r}
mpg[abs(z)>= 2 & abs(z) <= 3]
```
Using z, the values of mpg that is defined as outliers: 
```{r}
mpg[abs(z) > 3]
```
Dotplot using lattice package.
```{r}
mycol = ifelse(abs(z)>3, "Red",
               ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black")) 
mypch = ifelse(abs(z)>3, 20,
               ifelse(abs(z)>=2 &abs(z)<=3,21, 22))
library(lattice)
dotplot(mpg,col=mycol, main = "Plot of mpg values")
```


## Task 4

```{r}
boxplot(mpg, main = "Boxplot of mpg values", xlab = "MPG", col = "black", horizontal = TRUE, notch = TRUE)
```
Using Chebyshev's Theorem, at least 
 1 - 1/k^2 of the data lie within k standard deviations of the mean. 
 
 When k = 2,

```{r}
k = 2
chebval  = 1-1/(k^2)
chebval
```
The exact proportion of data within 2 standard deviation of the mean: 
```{r}
exactProportion = length(mpg[abs(z) < 2])/length(mpg)
exactProportion
```

In Chebyshev's Theorem, at least 8/9 of the data lie within 3 standard deviation of the mean. 
```{r}
k = 3
chebval = 1- 1/(k^2)
chebval
```

The exact proportion of data within 3 standard deviations of the mean: 
```{r}
exactProportion = length(mpg[abs(z)<3])/length(mpg)
exactProportion
```
Does Chebyshev agree with the data?

```{r}
compare = length(mpg[mean(mpg)-2*sd(mpg) < mpg & mpg < mean(mpg)+ 2*sd(mpg) ])/length(mpg)
compare
```

Yes, the value is approximately close to 8/9 of the data. Therefore, we can conclude that the Chebyshev agree with the data. 

The Chebyshev correspond well with the Empirical rule when the standard deviation is 2. The prediction of the amount of data within 2 standard deviation of the mean is around 96%. 

Therefore the Empirical Rule is quite valid in this case because the data is approximately bell-shaped. 














